Water discharge has a variable impact on sediment transport capacity in subglacial systems

### 2023-08-24 Delaneyetal2023b.tex
Small bits in discussion

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h). Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data in plots are at $15$\unit{min} intervals.
    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
    Note that the x and y axis are different for \alpine{} and \icesheet{} such that they each cover a summer.
    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilising and transporting sediment that subaerial ones}
  } 
  \label{fig:model_outs}
\end{figure}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

@@

The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge in to the Kangerlussuaq fjord \citep{mcgrath2010}.
This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
In these catchments, a less consistent relationships may persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}, possibly due to the variable link between sediment transport capacity and water discharge in subglacial conditions.

Second, the sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
The threshold of motion for sediment can be reached more frequently, and across a range of water discharges, in subglacial systems.
Therefore, glaciers could well  persist in a supply-limited regime, where sediment discharge also scales with sediment production from glacier quarrying and abrasion, along with melt water access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
Note, however, that the absences of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed  \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
\ian{might need to tighten this better}

>>>>

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs resulting from the hydrographs shown in panels a and e for the cases called \alpine (a-d) and \icesheet{} (e-h). Gray (black) lines in a and e represent the subglacial channel size (water discharge).  In other panels, blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data in plots are at $15$\unit{min} intervals.
    Arrows denote some events where the variable peaks in the subglacial channel prior to the subaerial one.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
    Note that the x and y axes are different for \alpine{} and \icesheet{} such that they each cover a summer.
    \ian{Andrew's comment: for a given channel width, subglacial channels are more effective at mobilizing and transporting sediment than subaerial ones}
  } 
  \label{fig:model_outs}
\end{figure}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks. Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

@@

The strong link between sediment discharge capacity and open channel flow conditions could be manifested ~ $20$\,\unit{km} downstream of the Leverett measurement site, where a strong correlation persists between sediment plume size and water discharge into the Kangerlussuaq fjord \citep{mcgrath2010}.
This lies in contrast to other marine-terminating glacier fjords in Greenland where the subglacial channels are pressurized by ocean water at the ice front \citep[e.g.][]{how2017}. 
In these catchments, a less consistent relationships may persist between water discharge, or melt extend, and sediment plume size \citep{tedstone2012}, possibly due to the variable link between sediment transport capacity and water discharge in subglacial conditions.

Second, the sediment discharge records could represent factors such as increased sediment accessibility when melt extends up glacier, thereby concurrently increasing sediment discharge with water discharge \citep[e.g.][]{delaney2020}.
In this case, the sediment transport records presented here may represent supply-limited sediment transport, less dependent on sediment transport capacity.
Glaciers could well trend toward a supply-limited regime, because the threshold of motion for sediment can be reached more frequently, and across a range of water discharges, compared to subaerial systems.
In a supply-limited system, sediment discharge would also scale with sediment production from glacier quarrying and abrasion, along with meltwater access \citep[e.g.][]{herman2015,ugelvig2018,delaney2023}.
 Sediment exhaustion through this process may also explain the stronger dependence of sediment discharge from the Greenland Ice Sheet on basal shear stress, a proxy for bedrock erosion, rather than glacier melt \citep{overeem2017}.
Note, however, that the absence of lags between sediment and water discharge may suggest that sediment is mobilized close to the measurement stations, as opposed to distal areas of the glacier bed as in the supply-limited regime \citep[Figure~\ref{fig:model_outs}\,a and e][]{williams1989}.
